Introduction to Markov state modeling with the PyEMMA software --- v\versionnumber

### 2018-07-26 manuscript.tex
resolve TODOs related to macrostate numbers

\begin{figure}
\includegraphics{figure_3}
\caption{Exemplary analysis of the conformational dynamics of a pentapeptide backbone:
(a) The reweighted free energy surface projected onto the first two independent components exhibits five minima which
(b) PCCA++ identifies as five metastable states.
(c) TODO:~The second right eigenvector shows that the slowest process shifts probability between the least probable state ($\mathcal{S}_1$) and the two most probable states ($\mathcal{S}_3$, $\mathcal{S}_4$), whereas
(d) the committor $\mathcal{S}_2\to\mathcal{S}_3$ indicates that the most probable state $\mathcal{S}_4$ acts as a transition state for transitions between states $\mathcal{S}_2$ and $\mathcal{S}_3$.
(e) The Trp-1 SASA shows high variance for all macrostates except $\mathcal{S}_1$ in which the Trp-1 shows the lowest exposure to the solvent.
(f) The Trp-1 SASA autocorrelation function yields a weak signal (top) which, however, can be enhanced if the system is prepared in the nonequilibrium condition $\mathcal{S}_1$ (bottom).
(g) Superimposed cartoon representations of $200$ molecular conformations sampled from each of the four macrostates show that the backbone appears to obstruct the Trp-1 exposure to the solvent in $\mathcal{S}_{(1,2)}$ but less so in $\mathcal{S}_{(3,4)}$.}
\label{fig:msm-analysis}
\end{figure}

@@

A spectral clustering using the PCCA++ algorithm~\cite{pcca++} allows us to coarse-grain the $75$ $k$-means microstates into five metastable macrostates (Fig.~\ref{fig:msm-analysis}b) $\mathcal{S}_i$, $i=1,\dots,5$, for which we then approximate the stationary probabilities and relative free energies (defined up to an additive constant)
\[ \begin{array}{ccc}
\textrm{macrostate } \mathcal{S}_i & \pi_{\mathcal{S}_i} & G_{\mathcal{S}_i} / \textrm{k}_\textrm{B} T \\
\hline
\mathcal{S}_1 & 0.004 & 5.565 \\
\mathcal{S}_2 & 0.013 & 4.314 \\
\mathcal{S}_3 & 0.021 & 3.864 \\
\mathcal{S}_4 & 0.962 & 0.039 \\
\mathcal{S}_5 & \text{TODO} & \text{TODO} \\
\end{array}\]
using the relation
\begin{equation}
\label{eq:fe}
G_{\mathcal{S}_i} = - \textrm{k}_\textrm{B} T \ln \sum\limits_{j\in \mathcal{S}_i} \pi_j,
\end{equation}
where $\pi_j$ denotes the MSM stationary weight of the $j^\textrm{th}$ microstate.

In order to interpret the slowest relaxation timescales, we refer to the (right) eigenvectors of the MSM as they contain information about what configurational changes are happening on which timescales.
The first right eigenvector corresponds to the stationary process and its eigenvalue is the Perron eigenvalue $1$.
The second right eigenvector, however, corresponds to the slowest process 
(the eigenvector components are real because of the detailed balance constraint enforced during MSM estimation).
The minimal and maximal components of the second right eigenvector indicate the microstates between which the process shifts probability density.
The relaxation timescale of this exchange process is exactly the corresponding implied timescale, which can be computed from its corresponding eigenvalue using~\eqref{eq:its}.
TODO:~In the projection onto the first two TICA components, we identify the slowest MSM process as a probability shift between macrostate $\mathcal{S}_1$ and the rest of the system, with macrostates $\mathcal{S}_3$ and $\mathcal{S}_4$ in particular (Fig.~\ref{fig:msm-analysis}c).

TODO:~The mean first passage times (MFPTs) out of and into the macrostate $\mathcal{S}_1$ compute to
\[ \begin{array}{crcr}
\textrm{direction} & \textrm{mean / ns} && \textrm{std / ns} \\
\hline
\mathcal{S}_1 \to \mathcal{S}_{(2,3,4)} & 8.9 & \pm & 1.8 \\
\mathcal{S}_{(2,3,4)} \to \mathcal{S}_1 & 2487.5 & \pm &  504.7
\end{array}\]
using the Bayesian MSM.

TPT~\cite{weinan-tpt,metzner-msm-tpt} is a method used to analyze the statistics of transition pathways. The TPT version of~\cite{noe-folding-pathways} can be conveniently applied to the estimated MSM. TODO:~Here, we compute the TPT flux between macrostates $\mathcal{S}_2$ and $\mathcal{S}_3$ (Fig.~\ref{fig:msm-analysis}d).
The committor projection onto the first two TICA components shows that it is constant within the metastable states defined above.
Transition regions (macrostate $\mathcal{S}_4$) can be identified by committor values $\approx \frac{1}{2}$.

>>>>

\begin{figure}
\includegraphics{figure_3}
\caption{Exemplary analysis of the conformational dynamics of a pentapeptide backbone:
(a) The reweighted free energy surface projected onto the first two independent components exhibits five minima which
(b) PCCA++ identifies as five metastable states.
(c) The second right eigenvector shows that the slowest process shifts probability between the least probable state ($\mathcal{S}_1$) and the other states, in particular states ($\mathcal{S}_4$, $\mathcal{S}_5$), whereas
(d) the committor $\mathcal{S}_2\to\mathcal{S}_4$ indicates that the most probable state $\mathcal{S}_5$ acts as a transition state for transitions between states $\mathcal{S}_2$ and $\mathcal{S}_4$.
(e) The Trp-1 SASA shows high variance for all macrostates except $\mathcal{S}_1$ in which the Trp-1 shows the lowest exposure to the solvent.
(f) The Trp-1 SASA autocorrelation function yields a weak signal (top) which, however, can be enhanced if the system is prepared in the nonequilibrium condition $\mathcal{S}_1$ (bottom).
(g) Superimposed cartoon representations of $200$ molecular conformations sampled from each of the four macrostates show that the backbone appears to obstruct the Trp-1 exposure to the solvent in $\mathcal{S}_{(1,2)}$ but less so in $\mathcal{S}_{(3,4)}$.}
\label{fig:msm-analysis}
\end{figure}

@@

A spectral clustering using the PCCA++ algorithm~\cite{pcca++} allows us to coarse-grain the $75$ $k$-means microstates into five metastable macrostates (Fig.~\ref{fig:msm-analysis}b) $\mathcal{S}_i$, $i=1,\dots,5$, for which we then approximate the stationary probabilities and relative free energies (defined up to an additive constant)
\[ \begin{array}{ccc}
\textrm{macrostate } \mathcal{S}_i & \pi_{\mathcal{S}_i} & G_{\mathcal{S}_i} / \textrm{k}_\textrm{B} T \\
\hline
\mathcal{S}_1 & 0.004 & 5.567 \\
\mathcal{S}_2 & 0.014 & 4.293 \\
\mathcal{S}_3 & 0.021 & 3.841 \\
\mathcal{S}_4 & 0.021 & 3.875 \\
\mathcal{S}_5 & 0.940 & 0.062 \\
\end{array}\]
using the relation
\begin{equation}
\label{eq:fe}
G_{\mathcal{S}_i} = - \textrm{k}_\textrm{B} T \ln \sum\limits_{j\in \mathcal{S}_i} \pi_j,
\end{equation}
where $\pi_j$ denotes the MSM stationary weight of the $j^\textrm{th}$ microstate.

In order to interpret the slowest relaxation timescales, we refer to the (right) eigenvectors of the MSM as they contain information about what configurational changes are happening on which timescales.
The first right eigenvector corresponds to the stationary process and its eigenvalue is the Perron eigenvalue $1$.
The second right eigenvector, however, corresponds to the slowest process 
(the eigenvector components are real because of the detailed balance constraint enforced during MSM estimation).
The minimal and maximal components of the second right eigenvector indicate the microstates between which the process shifts probability density.
The relaxation timescale of this exchange process is exactly the corresponding implied timescale, which can be computed from its corresponding eigenvalue using~\eqref{eq:its}.
In the projection onto the first two TICA components, we identify the slowest MSM process as a probability shift between macrostate $\mathcal{S}_1$ and the rest of the system, with macrostates $\mathcal{S}_4$ and $\mathcal{S}_5$ in particular (Fig.~\ref{fig:msm-analysis}c).

The mean first passage times (MFPTs) out of and into the macrostate $\mathcal{S}_1$ compute to
\[ \begin{array}{crcr}
\textrm{direction} & \textrm{mean / ns} && \textrm{std / ns} \\
\hline
\mathcal{S}_1 \to \mathcal{S}_{(2,3,4,5)} & 9.0 & \pm & 1.9 \\
\mathcal{S}_{(2,3,4,5)} \to \mathcal{S}_1 & 2496.4 & \pm &  470.0
\end{array}\]
using the Bayesian MSM.

TPT~\cite{weinan-tpt,metzner-msm-tpt} is a method used to analyze the statistics of transition pathways. The TPT version of~\cite{noe-folding-pathways} can be conveniently applied to the estimated MSM. Here, we compute the TPT flux between macrostates $\mathcal{S}_2$ and $\mathcal{S}_4$ (Fig.~\ref{fig:msm-analysis}d).
The committor projection onto the first two TICA components shows that it is constant within the metastable states defined above.
Transition regions (macrostate $\mathcal{S}_4$) can be identified by committor values $\approx \frac{1}{2}$.
